\documentclass[usegeometry,parskip=half]{scrartcl}

\usepackage[
  a4paper,
  top=25mm,
  bottom=25mm,
  left=30mm,
  right=30mm,
  footskip=10mm
]{geometry}

\usepackage[hidelinks]{hyperref}
\pagenumbering{gobble}

\usepackage{microtype}
\usepackage{fontspec}
\usepackage{unicode-math}

\setmainfont{Stix Two Text}
\setsansfont{TeX Gyre Heros}
\setmathfont{Stix Two Math} 

\begin{document}
\textbf{ICS0026 Cryptography --- evaluation criteria}

To pass, you must satisfy all three points:
\begin{enumerate}
  \item You must submit and pass each homework assignment.
  \begin{itemize}
    \item There will be at most $7$ homework assignments.
    \item You will have at least $2$ weeks to submit each homework.
    \item If you cannot meet the deadline, write to me no later than 4 days before the deadline with your reason.
    No extensions granted if you miss the notice period.
    \item When you are stuck:
    \begin{itemize}
      \item You may turn to me for help after classes/by email/by Discord.
      \item You may give pointers to each other, but not share solutions.
      \item If you cannot solve the practical part, it is enough to write about the technical issues you had, and show what approaches you tried.
    \end{itemize}
    \item Submitting an incomplete homework without explaining approaches you tried and asking me for tips $\implies$ failed homework $\implies$ failed course.
    \item If I have doubts about your homework (e.g. suspect plagiarism), I will ask you extra questions as a second chance.
  \end{itemize}
  \item Attend 8 lectures and 8 practices OR pass the pre-exam.
  \begin{itemize}
    \item Pre-exam not required if you have attended at least 8 lectures and 8 practices.
    \item 30min closed book with basic questions (easier than the exam) to make sure you have not ghosted the course.
    \item If you fail: we can discuss the results to make up for some lost points, but there are no retakes.
    Consider it potential extra credit for lost attendance points.
  \end{itemize}
  \item You must attend and pass the in-person written exam.
  \begin{itemize}
    \item 1h 30min closed book exam with a handwritten one-sided A4 cheat-sheet allowed.
    \item Alternatively: 20min oral exam (max 10 students).
    \item $1$ retake allowed.
    \item I will publish an exam help/topics guide.
    \item All previous assignments must be passed to attend.
  \end{itemize}
\end{enumerate}

The reasons for the above criteria and for the closed book exam are explained in the intro lecture.

\vspace*{\parskip}

Violation of good academic practice:

Copying someone's homework, plagiarism, etc.\footnotemark{}, fails you the course and I will report you.

\footnotetext{Chapter 9 of TalTech's academic policies: \url{https://taltech.ee/en/academic-policies-taltech}}
\end{document}
