\documentclass{practice}

\title{4}
\date{\today}

\usepackage{crysymb}

%\usepackage{etoolbox}
%\makeatletter
%\preto{\@verbatim}{\topsep=0pt \partopsep=-2\parsep }
%\makeatother

\fvset{listparameters=\setlength{\topsep}{0pt}\setlength{\partopsep}{0pt}}

\usepackage{listings}

\lstset{
basicstyle=\small\ttfamily,
columns=flexible,
breaklines=true
}

\begin{document}
\maketitle

\begin{task}{Modulo bias}
  \textit{Preface.}
  A common task in programming and especially in public-key cryptography is to generate a random number in a range.

  In cryptography, it is crucial that random numbers are sampled uniformly at random, meaning that the probability of picking any number from a set/range of distinct elements is the same.
  For example, by sampling uniformly at random from the set $\{0, 1, 2, 3, 4\}$, i.e. $x\getsu\{0, 1, 2, 3, 4\}$, the probability distribution should be the following:
  \begin{center}
    \begin{tabular}{l r r r r r}
      \toprule
      $x$ & $0$ & $1$ & $2$ & $3$ & $4$\\
      \midrule
      $\Pr$ & $0.2$ & $0.2$ & $0.2$ & $0.2$ & $0.2$\\
      \bottomrule
    \end{tabular}
  \end{center}

  During week 2, we saw that we could get roughly those results by querying random bytes from the CSPRNG.
  However, querying even a single byte of \enquote{randomness} would give us an integer in the range $0$--$255$.

  \iffalse
  In general, getting a random number in the range $[0, 2^n - 1]$ is `easy' since then we could just bit-mask the bytes gotten from the CSPRNG.
  For example, if we want a random number in the range $[0, 7]$, we simply need to isolate the first three bits of a random byte: \texttt{0b111} is $7$.
  \fi

  So, how could we get a number in the range $[0, 4]$?
  The naive approach is to make use of the \emph{modulo} operator: $a \bmod n$ gives us a result in the set $\{0, 1, \dots, n-1\}$ for two integers $a, n$.
  However, this approach can actually skew the probability distribution.
  The following table shows the number of occurrences of numbers $0$--$5$ for $a \bmod n$ where each $a\in\{0, \dots, 255\}$ is picked once and $n = 5$:
  \begin{center}
    \begin{tabular}{l r r r r r}
      \toprule
      $a \bmod 5$ & $0$ & $1$ & $2$ & $3$ & $4$\\
      \midrule
      Num. of occ. & $52$ & $51$ & $51$ & $51$ & $51$\\
      \bottomrule
    \end{tabular}
  \end{center}

  We see that the outcome $0$ occurs one more time than the other values, which may not seem like much, but clearly indicates that the outcome distribution is no longer uniform.

  \textit{Task.}
  Your task is to generate ten million random $\bigl(10^7\bigr)$ $8$-bit integers (using the \texttt{secrets} module) and generate and plot (as a bar chart) the frequency distributions for the following modulus values: $100$, $150$, $200$.
  Observe how the bias shifts.

  Now do the same, except generate the random values within the ranges defined by the three moduli using simple rejection sampling (regenerate values until they fit in the range).
  Do the distributions look better?

  You could use Matplotlib\footnote{\url{https://matplotlib.org}} and pandas\footnote{\url{https://pandas.pydata.org}} for generating the charts/visualisations.

  \begin{tcolorbox}[title=Good practice]
    In practice, to get cryptographic random numbers in a range, you should use the functions provided by a reliable cryptographic library instead of implementing your own functionality.
    \tcblower
    For Python, you could use 
    \href{https://docs.python.org/3/library/secrets.html#secrets.randbelow}{\texttt{secrets.randbelow()}}.
  \end{tcolorbox}
\end{task}

\begin{task}{MP arithmetic}
  \textit{Preface.}
  If you have programming background in strongly typed languages (e.g. C/C++, Java/C\#) you likely know that you are typically limited to $64$-bit integers (though $128$-bit integer support is becoming more common in C/C++) for built-in types.
  Without getting into the specifics of how processors work, we can consider the following simplification: a $64$-bit CPU implements operations (addition, multiplication, etc.) for $64$-bit numbers.

  In class we saw that to get $128$-bits of security against classical computers (i.e. not quantum computers), we should use $3072$-bit RSA, that is, the RSA modulus (semiprime) should be a $3072$-bit integer.
  So how can we perform mathematical operations on such numbers if they do not fit into our CPU registers, and the operations are not provided by the processor?
  The answer is to use software-implementations, that is, we emulate the storage and computation in software.
  This is typically known as \emph{arbitrary-precision} or \emph{multiple-precision (MP) arithmetic}.

  Python is not a strongly typed language, and takes care of implementing MP arithmetic for us.
  This is why we typically need not worry about overflow in Python when working with numbers.
  While it is not surprising that MP arithmetic is slower than \enquote{hardware arithmetic}, then not all MP implementations are created equal.
  That is, when doing heavy number crunching, we may want better performance than what Python provides us by default.

  One of the most popular and performant general purpose MP libraries is the GNU Multiple Precision library (GMP)\footnote{\url{https://gmplib.org}}.
  To get enhanced performance for MP computations in Python, the \href{https://pypi.org/project/gmpy2/}{\texttt{gmpy2}} module is a neat pick.
  Depending on your setup (mainly the OS) it might be slightly tricky to get it to work.
  If you cannot get it to work on Windows, I recommend attempting the tasks on a Linux VM unless you feel comfortable troubleshooting the issues yourself.

  GMP offers way more than simply enhanced performance.
  For example, it implements different primality testing algorithms and other functions that make working with primes more convenient for us in a learning setting.

  \textit{Task.}
  Your task is to generate three $3072$-bit numbers and then perform the modular exponentiation $a^b \bmod c$ such that $a < b < c$, with $c$ odd.

  Perform this exponentiation $10$ times with Python's built-in \href{https://docs.python.org/3/library/functions.html#pow}{\texttt{pow()}} function, $10$ times with the \href{https://gmpy2.readthedocs.io/en/latest/mpz.html#gmpy2.powmod}{\texttt{gmpy2.powmod()}} function, and $10$ times with the \href{https://gmpy2.readthedocs.io/en/latest/mpz.html#gmpy2.powmod_sec}{\texttt{gmpy2.powmod\_sec()}} function.
  Note that \texttt{powmod\_sec()} requires the modulus to be odd (can you think of a reason?).
  What do you think about the relative speed of these functions?

  Measure the time it takes for each of the $10$ loops and compare the results, e.g. with \href{https://docs.python.org/3/library/time.html#time.perf_counter}{\texttt{time.perf\_counter()}}.
  Are the results consistent with your guess?

  Finally, try the naive implementation, i.e. \texttt{a ** b \% c}.
  Let it run for a few seconds, then abort the program execution---you have a life to live.

  \begin{tcolorbox}[title=Warning]
    General purpose MP libraries are not designed with cryptography in mind, rather they are meant to be as efficient as possible.
    This means that they are riddled with different side-channel leaks (among other issues) and are not suited for cryptographic uses where side-channel resistance is needed (e.g. decryption).
    \tcblower
    While GMP does provide functions with some side-channel protections\footnote{GMP developers' secure corner: \url{https://gmplib.org/devel/sec}}, you should not consider these as a substitution for cryptographic libraries.
  \end{tcolorbox}
\end{task}

\begin{task}{\enquote{Generating} random primes}
  \textit{Preface.}
  Primes and primality testing are awesome and intriguing topics.
  However, they are also quite mathematically involved, so we will not go into the details in this course.
  Still, there are some things you should know about primality testing.

  There are various \emph{primality testing algorithms} to test whether a generated number is prime.
  For large numbers, primality tests are typically \emph{probabilistic}: if a number is found to be composite, then it is so, and if not, the number is likely a prime, but not provably so.
  In general, for robust primality testing, either or both of the following approaches are used:
  \begin{itemize}
    \item Run multiple iterations of a \enquote{quality} primality test.
    Intuitively, the more iterations a number passes, the more likely it is for the number to be prime.
    \item Combine different primality tests.
  \end{itemize}

  Perhaps the most widely used probabilistic primality testing algorithm is the \emph{Miller-Rabin}\footnote{\url{https://en.wikipedia.org/wiki/Miller–Rabin_primality_test}} test.
  
  The question then is: how to pick the number of iterations for the Miller-Rabin test?
  The rule of thumb is that Miller-Rabin gives a \enquote{security level} of $128$-bits for $64$ iterations, but there are different methods to select the number of \enquote{optimal} iterations.
  For example, OpenSSL uses the Miller-Rabin test with $128$ iterations for integers larger than $2048$ bits\footnote{\url{https://github.com/openssl/openssl/blob/master/crypto/bn/bn_prime.c\#L94}}.

  There are different approaches for generating (finding) random primes, but the two most intuitive ones are:
  \begin{enumerate}
    \item \emph{Random search}: keep selecting a random $n$-bit number until it passes some primality test.
    \item \emph{Incremental search}: instead of rejection sampling, keep incrementing the generated number by $2$ until it passes some primality test.
  \end{enumerate}

  If it seems to you that the second method is \enquote{less secure}, your intuition is quite good: indeed the latter method introduces some bias not present with the first method.
  However, currently no (publicly) known method can help an attacker abuse this bias.

  \textit{Task.}
  Generate two $1536$-bit primes using \emph{random search}.
  Benchmark the time it took to find the two primes.
  Now, count the number of bits in the product of these numbers.
  Is it $3072$?

  Next, implement a simple \emph{incremental search} for finding two $1536$-bit primes and benchmark the time taken.
  Count also the number of bits in the semiprime product.

  Finally, use a library function for finding two $1536$-bit primes and benchmark the time taken.
  Once more, count the number of bits in the semiprime.

  Compare the performance of all three methods.

  \textit{Hint.}
  The following functions are likely to help you:
  \begin{itemize}
    \item \href{https://gmpy2.readthedocs.io/en/latest/mpz.html#gmpy2.mpz.is_prime}{\texttt{gmpy2.is\_prime()}} and \href{https://gmpy2.readthedocs.io/en/latest/mpz.html#gmpy2.mpz.is_probab_prime}{\texttt{gmpy2.is\_probab\_prime()}}
    \item \href{https://gmpy2.readthedocs.io/en/latest/mpz.html#gmpy2.next_prime}{\texttt{gmpy2.next\_prime()}}
    \item \href{https://docs.python.org/3/library/stdtypes.html#int.bit_length}{\texttt{int.bit\_length()}} and \href{https://gmpy2.readthedocs.io/en/latest/mpz.html#gmpy2.bit_length}{\texttt{gmpy2.bit\_length()}}
  \end{itemize}

  Check out also gmpy2's docs on \href{https://gmpy2.readthedocs.io/en/latest/advmpz.html#advanced-number-theory-functions}{\enquote{Advanced Number Theory Functions}} for more granular control on primality testing.

  \begin{tcolorbox}[title=Warning]
    The methods we use here are for learning purposes/convenience only.
    There are many parameters that go into picking cryptographic primes which we have not covered and will not cover.
    You must not attempt to roll your own cryptography/generate your own keys this way.
    \tcblower
    If you want to get an idea of the complexity of picking suitable primes, read appendix \href{https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.186-5.pdf#page=42}{A.1} of \href{https://csrc.nist.gov/pubs/fips/186-5/final}{FIPS 186-5}.
  \end{tcolorbox}
\end{task}

\newpage

\begin{task}{Sage RSA}
  SageMath\footnote{\url{https://www.sagemath.org}} is a computer algebra system (CAS) which can be programmed in with a Python-based language.
  It is free and open-source as opposed to Wolfram Mathematica and MATLAB, and is very powerful.
  As such, it is quite popular among cryptographers for some quick computations/prototyping, and is also popular for cryptography tasks in CTFs (Capture The Flag exercises).

  Check out also Sage's \href{https://doc.sagemath.org/html/en/reference/cryptography/index.html}{\enquote{cryptography reference}}.

  \textit{Task.}
  Given the following RSA public key:
  \begin{itemize}
    \item $e = 65537$
    \item $n = 230624063920680065513680711198246109591$,
  \end{itemize}
  decipher the RSA ciphertext $c = 161670843286605628968182075437059446607$.

  Note that the RSA modulus is $128$ bits long, so \enquote{naively} brute forcing either prime is not a good approach.
  Can you think of a more efficient approach?
  
  \textit{Hint 1.}
  RSA-100 ($330$ bits) was \emph{factored} way back in 1991.

  \textit{Hint 2.}
  Recovering the message takes less than a second on my laptop with the proper Sage tools.

  \textit{Addendum.}
  Check out the RsaCtfTool\footnote{\url{https://github.com/RsaCtfTool/RsaCtfTool}}.
  Can you solve the task using the tool?
\end{task}

\begin{task}{OpenSSL \& PEM}
  \textit{Preface.}
  Last week we learnt a bit about OpenSSL and we used it for AES encryption and decryption.
  OpenSSL has a ton of functionality, and among this functionality is the capability to generate RSA keys and perform RSA encryption/decryption.

  When generating RSA keys with OpenSSL, you may notice that by default, the key material is base64-encoded and sandwiched between the header and footer:
  \begin{Verbatim}
-----BEGIN PRIVATE KEY-----
...
-----END PRIVATE KEY-----
  \end{Verbatim}
  or
  \begin{Verbatim}
-----BEGIN PUBLIC KEY-----
...
-----END PUBLIC KEY-----
  \end{Verbatim}
  This format is called the PEM (Privacy-Enhanced Mail) format\footnote{\url{https://en.wikipedia.org/wiki/Privacy-Enhanced_Mail}} and the convention for key-files to have the extension \texttt{.pem} or \texttt{.key}.
  We will see other headers and usages of PEM during future classes/practices.

  There are three primary OpenSSL tools for working with public keys:
  \begin{itemize}
    \item \href{https://docs.openssl.org/master/man1/openssl-genpkey/}{\texttt{genpkey}}: private key generation (and more)
    \item \href{https://docs.openssl.org/master/man1/openssl-pkey/}{\texttt{pkey}}: commands to \enquote{process} keys, e.g. change their format (PEM vs DER) or their passphrase
    \item \href{https://docs.openssl.org/master/man1/openssl-pkeyutl/}{\texttt{pkeyutl}}: the utility for encrypting and decrypting (also signing and verifying signatures)
  \end{itemize}

  \textit{Task.}
  For this task, you will generate an RSA keypair and use it to encrypt and decrypt data:
  \begin{enumerate}
    \item Generate an RSA private key using OpenSSL with
    \begin{Verbatim}
openssl genpkey -algorithm RSA -out {path_private_pem}
    \end{Verbatim}
    What is the default bit-size for the key?

    \begin{tcolorbox}[title=Warning]
      By generating a key using this command, the key is \emph{not} protected by a passphrase, meaning that anyone with access to the key-file/key can simply use it for decryption.
      \tcblower
      Password protecting keyfiles is good practice (e.g. SSH or PGP keys).
      The following command generates and encrypts the RSA private key with AES-128 using the user-provided passphrase:
      \begin{Verbatim}
openssl genpkey -algorithm RSA -aes128
      \end{Verbatim}
    \end{tcolorbox}

    \item Generate a new 3072-bit RSA secret key and determine the public key corresponding to it.
    \item Finally, attempt to encrypt and decrypt files using the generated keys.
    Is there a limit to the file-sizes?
    Does OpenSSL use the proper RSA encryption padding by default?
  \end{enumerate}

  \textit{Note.}
  There exist also the utilities:
  \begin{itemize}
    \item \href{https://docs.openssl.org/master/man1/openssl-genrsa/}{\texttt{genrsa}}
    \item \href{https://docs.openssl.org/master/man1/openssl-rsa/}{\texttt{rsa}}
    \item \href{https://docs.openssl.org/master/man1/openssl-rsautl/}{\texttt{rsautl}}
  \end{itemize}
  which accomplish the same tasks but only for RSA.
  Therefore, these commands are a bit simpler.

  However, note that \texttt{rsautl} has been deprecated and \texttt{pkeyutl} should be used instead.
  More generally, you should prefer the generic commands since that seems to be the direction OpenSSL is moving in.
\end{task}

\begin{task}{ASN.1}
  \textit{Preface.}
  The base64 encoding is an encoding of the ASN.1 DER format which we briefly saw in class.
  A great CLI tool for visualising ASN.1 data structures is \texttt{dumpasn1}.

  \textit{Task.}
  Install \texttt{dumpasn1} and try to visualise what the key files really contain.
  Note that you must strip the PEM header and decode the base64 data before \texttt{dumpasn1} can parse the ASN.1 bytes.

  Now try to figure out how to use OpenSSL's \href{https://docs.openssl.org/master/man1/openssl-asn1parse/}{\texttt{asn1parse}} tool to do the same thing.
  Which output do you prefer?
\end{task}

\begin{task}{RSA in Python}
  \textit{Task.}
  Use a reliable 3\textsuperscript{rd} party library to implement RSA key generation, encryption, and decryption.
  Encryption should use OAEP padding, and the key-size should provide $128$-bits of security.

  Can you find test vectors similar to the NIST AES test vectors anywhere to test the library?
\end{task}

\end{document}
