\usetheme{Shamir}

\usepackage{graphicx}

\usepackage{csquotes}
\setquotestyle{british}

\definecolor{oxfordblue}{RGB}{4,30,66}

\iffalse
\setbeamertemplate{headline}
{%
    \begin{picture}(0,0)
        \put(314,-50){%
            Owo?
        }
        \put(20,-55){%
            \rule{320pt}{0.4pt}
        }
    \end{picture}
}
\fi

\iffalse
\setbeamertemplate{footline}
{%
    \begin{picture}(0,0)
        \put(20,30){%
            \rule{320pt}{0.4pt}
        }
        \put(20,14){%
            Hihi
        }
        \put(100,14){%
            \color{oxfordblue}\insertshortdate
        }
        \put(160,14){%
            \color{oxfordblue}\insertshorttitle
        }
        \put(337,14){%
            \color{oxfordblue}\insertpagenumber/\inserttotalframenumber
        }
    \end{picture}%
}
\fi

\title[Short version of title]{ICS0026 Cryptography}
\subtitle{Course organisation}
\author%
{%
    \sc{Taaniel Kraavi}\\
    \textit{\footnotesize Guest Lecturer}
}
\institute%
{%
    \textit{IT College}\\
    \textit{Tallinn University of Technology}
}

\begin{document}
\begin{frame}[plain]
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Course aim}

  \begin{columns}[T]
    \begin{column}{.4\textwidth}
      \textbf{Prerequisites:}
      \begin{itemize}[<+->]
        \pause
        \item Know basic programming
        \begin{itemize}
          \item C \& python3
        \end{itemize}
        \item Notions in:
        \begin{itemize}
          \item Discrete mathematics
          \item Binary data representation
          \item Group theory (bonus)
        \end{itemize}
      \end{itemize}
    \end{column}

    \pause
    \begin{column}{.4\textwidth}
      \textbf{Approach:}
      \begin{itemize}[<+->]
        \item Common terminology
        \item Cryptographic primitives as a black-box API
        \item Pitfalls
        \item DYOR: due diligence
      \end{itemize}
    \end{column}
  \end{columns}

  \vspace*{2em}

  \pause
  ITC8240 (master's course) for mathematical cryptography.
\end{frame}

\begin{frame}
  \frametitle{Schedule}

  \textbf{Lectures}

  Tuesdays: 14:00--15:30 room ICO-316

  \vspace*{1em}

  \pause
  \textbf{Practices}

  Thursdays: 14:00--15:30 room ICO-319

  \vspace*{1em}

  \pause
  Mandatory to attend at least 8 lectures and 8 practices.

  \pause
  Valid exceptions:
  \begin{itemize}
    \item work
    \item scheduling conflicts
  \end{itemize}

  \pause
  Exceptions $\implies$ 20min oral exam.
\end{frame}

\begin{frame}
  \frametitle{Grading}

  The course is pass/fail.

  \vspace*{1em}

  \pause
  \textbf{Homework}
  \begin{itemize}[<+->]
    \item Up to 8 mandatory assignments
    \item 2 weeks to complete
    \item Theoretical part (research) \& practical part (programming)
    \item Best-effort:
    \begin{itemize}
      \item Important to try
      \item Explain your issues
    \end{itemize}
    \item Semi-random defences
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Grading}

  The course is pass/fail.

  \vspace*{1em}

  \textbf{Exam}
  \begin{itemize}[<+->]
    \item Mandatory
    \item Pre-report ($\le$1 page) on knowledge gaps
    \item 5--10min oral assessment:
    \begin{itemize}
      \item questions about basic concepts (closed book)
      \item report discussion
    \end{itemize}
    \item +10min if you skipped classes
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Grading}

  The course is pass/fail.

  \vspace*{1em}

  \textbf{Summary}
  \begin{itemize}
    \item All homework mandatory ($\le$8)
    \item $\ge$50\% attendance a bonus
    \item Report ($\le$1 page) on knowledge gaps
    \item 5--10min closed-book oral assessment
  \end{itemize}

  \vspace*{1em}

  \pause
  Not hard to pass so please do not offload homework to LLMs.
\end{frame}

\begin{frame}
  \frametitle{Syllabus}

  \begin{columns}[t]
    \begin{column}{.4\textwidth}
      \begin{enumerate}
        \item Introduction
        \item Symmetric cryptography
        \item Symmetric cryptography (cont.)
        \item Public-key cryptography
        \item Public-key cryptography (cont.)
        \item Hash functions
        \item MACs \& digital signatures
        \item Public-key infrastructure
      \end{enumerate}
    \end{column}

    \begin{column}{.4\textwidth}
      \begin{enumerate}\setcounter{enumi}{8}
        \item Authentication \& commitments
        \item TLS, HTTPS, SSH
        \item Secure programming \& hardware
        \item Zero-knowledge proofs
        \item Secret sharing \& MPC
        \item Internet voting
        \item Post-quantum cryptography
        \item Review
      \end{enumerate}
    \end{column}
  \end{columns}

  \vspace*{2em}

  \pause
  Topics might not be cleanly split by week.
\end{frame}

\begin{frame}
  \frametitle{Reading materials}

  The book \enquote{The Joy of Cryptography} by Mike Rosulek is freely available at \url{https://joyofcryptography.com} and covers many of the topics in this course.

  \vspace*{1em}

  The book is more mathematical than this course, but may hence provide you with valuable insight and a stronger understanding of the same concepts. 
\end{frame}

\end{document}
